---
title: "evaluate_SDMs"
author: "steppe"
date: "3/13/2022"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    includes:
      in_header: float_free.tex
  word_document: default

---

\tableofcontents 
\listoffigures
\listoftables

# Import and Wrangle Data

```{r load libraries, results='hide', message=F, warning=F}
library(tidyverse) # data tidying
library(here)
set.seed(12)

source(here::here('scripts/sdm_functions.R'))
```

Pull in the model evaluation data
```{r Import and wrangle data}
fnames <- paste0(paste0(here(), '/results/stats/'),
       list.files(paste0(here(), '/results/stats/'), pattern = '*glm')
)

data <- do.call("rbind", sapply(fnames, read.csv, simplify = FALSE)) %>% 
  rownames_to_column('Taxon') %>% 
  dplyr::select(1,3:6) %>% 
  mutate(Model = if_else(modelID %in% c(1,2,3), 'GLM', 'GAM')) %>% 
  mutate(Taxon = str_remove(Taxon, paste0(here(), '/results/stats/'))) %>% 
  mutate(Taxon = str_remove(Taxon, '[.]csv[.]\\d')) %>% 
  separate(Taxon, c('Taxon', 'Time'), '_glm_') %>% 
  mutate(Time = str_replace(Time, '_', ' ')) %>% 
  mutate(Time = as.POSIXct(Time, "%Y-%m-%d %H:%M:%S")) %>% 
  group_by(Taxon, Time) %>% 
  mutate(Event = cur_group_id()) %>% 
  ungroup() 

rm(fnames)
```

Create diff time objects to study how long it took to produce each ensemble forecast. 
```{r}

data_diff_time <- data %>% 
  distinct(Taxon, Event, .keep_all = T) %>% 
  arrange(Time) %>% 
  mutate(Duration = as.numeric(difftime(Time, lag(Time), tz = "America/Chicago", units = "mins"))) %>% 
  mutate(Duration = ifelse(Duration >= 180, NA, Duration)) %>% 
  dplyr::select(Taxon, Event, Duration)

data <-left_join(data, data_diff_time, by = c('Taxon', 'Event'), copy = F)

rm(data_diff_time)
```

Pull in the data which were used to generate the models.
```{r}
PA_REG_DF_spatial <- sf::read_sf(here("data/processed/Pseudo_absences_regression.shp")) 

PA_REG_DF <- PA_REG_DF_spatial %>% 
  sf::st_drop_geometry() %>% 
  select(1:4) %>% 
  group_by(binomial) %>% 
  mutate(occurrence = ifelse(occurrence == 2, 0, occurrence)) %>% 
  mutate(Data_type = case_when(
    occurrence == 1 ~ 'Presence',
    occurrence == 0 & !is.na(PlotKey) ~ 'T_Absence',
    occurrence == 0 & is.na(PlotKey) ~ 'P_Absence'
  )) %>% 
  group_by(binomial, Data_type) %>% 
  summarise(Number_Occ = n()) %>% 
  pivot_wider(names_from = Data_type, values_from = Number_Occ) %>% 
  group_by(binomial) %>% 
  mutate(Prop_PA = sum(T_Absence, P_Absence)/Presence, # for each Presence how many absence
         Prop_PATA =  P_Absence / T_Absence, # for each true absence how many false
  ) %>% 
  pivot_longer(cols = P_Absence:T_Absence, names_to = 'occurrence_type', 
               values_to = 'occurrence_type_cnt')

data <-left_join(data, PA_REG_DF, by = c('Taxon' = 'binomial'), copy = F)

rm(PA_REG_DF)
```
Calculate three metrics of dispersion for the environmental presence data

```{r}
PA_REG_DF_spatial <- PA_REG_DF_spatial %>% 
  mutate(occurrence = ifelse(occurrence == 2, 0, occurrence)) %>% 
  filter(occurrence == 1)
spp_occurrence_REG_list <- split(PA_REG_DF_spatial, PA_REG_DF_spatial$binomial)

shout <- bind_rows(lapply(spp_occurrence_REG_list, variance_calcs))
```

Calculate the area occupied by the taxa within the study region

```{r}
areas_occupied <- bind_rows(lapply(spp_occurrence_REG_list, total_area)) %>% 
  mutate(area = as.numeric(area)/10)
data <- left_join(data, areas_occupied, by= c('Taxon' = 'binomial'))

rm(areas_occupied)
```



# Data Exploration


```{r}
GGally::ggpairs(data[, c(4:6, 10:11,13:14)])
```


```{r Relationship between evaluation metrics}
colors <- c("#56B4E9", "#E69F00")
colors <- colors[as.factor(data$Model)]
scatterplot3d::scatterplot3d(data[,4:6], pch = 5, color = colors,  grid=TRUE, type="p")
```

```{r Do seperate modelling runs give similar model evaluation statistics}

data %>% 
  distinct(Taxon, modelID, Event, .keep_all = T) %>% 
  select(Taxon, modelID, TSS, Event) %>% 
  group_by(Taxon) %>% 
  mutate(evaluation_mean = sum(Event)/n()) %>% 
  filter(evaluation_mean != Event) %>% 
  mutate(Group = ifelse(Event < evaluation_mean, 'A', 'B')) %>% 
  
  ggplot(aes(as.factor(Group), TSS))+
  geom_boxplot(notch = T, outlier.shape = 1) +
  geom_jitter(alpha = 0.5, width = 0.2) +
  theme_classic(base_size = 10) +
  labs(title = 'Comparision of Species which have had two ensembles created',
       y = 'True Skill Statistic (TSS)', x = 'Groups, 50 species replicated') +
  theme(plot.title = element_text(hjust = 0.5))

```


# Analyses

How long does it take to run an ensemble forecast? # THIS DOES NOT MAKE TOTAL SENSE AS TIME(GLM) =? TIME(GAM)

```{r Duration of time required to ensemble forecasts}

how_long <- data %>% 
  distinct(Taxon, Event, modelID, .keep_all = T)  %>% 
  select(Taxon, modelID, Model, Event, Duration) %>% 
  na.omit() %>% 
  group_by(Taxon, Event) %>% 
  mutate(Number_Models = n()) %>% 
  distinct(Taxon, Event, .keep_all = T)
  
ggplot(how_long, aes(as.factor(Number_Models), Duration))+
  geom_boxplot(notch = T, outlier.shape = 1) +
  geom_jitter(alpha = 0.5, width = 0.2) +
  theme_classic(base_size = 10) +
  labs(title = 'Time to complete an ensemble prediction',
       y = 'Duration Minutes', x = 'Number of Fitted Models') +
  theme(plot.title = element_text(hjust = 0.5))

fligner.test(Duration ~ as.factor(Number_Models), how_long)

how_long$Duration <- round(how_long$Duration, 0)
glm_poiss <- glm(Duration ~ Number_Models, family = poisson, data = how_long)
summary(glm_poiss)
AER::dispersiontest(glm_poiss,trafo=1)

summary(glm_negbin <- MASS::glm.nb(Duration ~ Number_Models, data = how_long))
pchisq(2 * (logLik(glm_negbin) - logLik(glm_negbin)), df = 1, lower.tail = FALSE)
(est <- cbind(Estimate = coef(glm_negbin), confint(glm_negbin)))
exp(est)

rm(how_long, glm_negbin, glm_poiss, est)
```

Fligner-Killeen test null hypothesis: the sample variances have equal variance, this is soundly rejected for the alternative. 

On average it looks like it may take 65% more time to ensemble a prediction as each additional model is added. 

Do the Number of Presence/Absences predict the accuracy of models ?

```{r}
data_pres <- data %>% filter(occurrence_type == 'Presence')
occurrence_pred_tss <- lm(TSS ~ occurrence_type_cnt, data = data)
occ_pr_tss_sum <- summary(occurrence_pred_tss)
occ_pr_tss_sum

ggplot(data_pres, aes(x = occurrence_type_cnt, y = TSS, color = Model))+
  geom_jitter()+
  #geom_smooth(method = "lm", formula = 'y ~ x') + 
  geom_abline(intercept = occ_pr_tss_sum$coefficients[1,1], 
              slope = occ_pr_tss_sum$coefficients[2,1], size = 2) +
  theme_classic(base_size = 10) +
  labs(title = 'Relationship between Number of Presence Records and TSS',
       y = 'TSS', x = 'Number of Presence') +
  theme(plot.title = element_text(hjust = 0.5))

rm(data_pres, occurrence_pred_tss, occ_pr_tss_sum)
```

Obviously more data does not make prediction worse, rather the root cause is likely to be the correlation between the number of records and the number of habitat types the species grows in. The more generalized a species, the more likely that a presence is to occur in an area deemed as predicted as absence. 


```{r}
data_pres <- data %>% filter(occurrence_type == 'Presence')
occurrence_pred_tss <- lm(TSS ~ Prop_PA, data = data)
occ_pr_tss_sum <- summary(occurrence_pred_tss)
occ_pr_tss_sum

ggplot(data_pres, aes(x = Prop_PA, y = TSS, color = Model))+
  geom_jitter()+
  #geom_smooth(method = "lm", formula = 'y ~ x') + 
  geom_abline(intercept = occ_pr_tss_sum$coefficients[1,1], 
              slope = occ_pr_tss_sum$coefficients[2,1], size = 2) +
  theme_classic(base_size = 10) +
  labs(title = 'Relationship between Number of Presence Records and TSS',
       y = 'TSS', x = 'Proportion of Presence records to Absence Records') +
  theme(plot.title = element_text(hjust = 0.5))
```


Do the accuracy of models decrease with increases in the geographic extent which the species ranger covers in the study area?

```{r}

ggplot(data, aes(x = area, y = TSS, color = Model))+
  geom_jitter()+
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  theme_classic(base_size = 10) +
  labs(title = 'Relationship between Area occupied by Taxon and TSS',
       y = 'TSS', x = 'Area occupied in km^2') +
  theme(plot.title = element_text(hjust = 0.5))

rm(areas_occ)
```

How does variance affect model fit? 
```{r}

variance_df <- data %>% 
  select(Taxon, modelID, TSS, Model) %>% 
  distinct(.keep_all = T) %>% 
  left_join(., shout, by = c('Taxon' = 'binomial'))
  
occurrence_pred_tss <- lm(TSS ~ Dispersion, data = variance_df)
occ_pr_tss_sum <- summary(occurrence_pred_tss)
occ_pr_tss_sum

variance_df %>% 
  filter(Dispersion == 'IQR') %>% 
  ggplot( aes(x = Value, y = TSS)) + 
  facet_wrap(~Dispersion) +
  geom_jitter() +
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  theme_classic(base_size = 10) +
  labs(title = 'Does variance predict TSS results?',
       y = 'TSS', x = 'Variance') +
  theme(plot.title = element_text(hjust = 0.5)) 

rm(variance_df, occurrence_pred_tss, occ_pr_tss_sum)
```

hmm not as expected, bigger sample size and more is explained, 

Let's try and fit a model

```{r}

ggplot(data, aes(round((-log(data$TSS))*100,0))) +
  geom_histogram()

glm_poiss <- glm(round((-log(data$TSS))*100,0) ~ data$area + data$occurrence_type_cnt, family = poisson, data = data)
summary(glm_poiss)
AER::dispersiontest(glm_poiss, trafo=1)

summary(glm_area <- glm(round((-log(TSS))*100,0) ~ area, family = poisson, data = data))
summary(glm_records <- glm(round((-log(TSS))*100,0) ~ occurrence_type_cnt, family = poisson, data = data))

AER::dispersiontest(glm_area, trafo=1)
AER::dispersiontest(glm_records, trafo=1)

modelTABLE <- MuMIn::model.sel(glm_area,glm_records)
modelTABLE

rm(glm_poiss, glm_area, glm_records)
```

