---
title: "evaluate_phenology_predictions"
author: "steppe"
date: '2022-04-23'
output: html_document
---


```{r load libraries, results='hide', message=F, warning=F}
library(here)
library(lubridate)
library(plotrix)
source(here::here('scripts/phen_functions.R'))
library(tidyverse) #data tidying
```

We can only test the accuracy of the Weibull method on a subset of cerca 60 species we have created estimates for. An interesting component of the Weibull development paper is that the authors did not actually attempt to estimate phenological events using herbarium records and then test them using the RMBL data set, but rather used the same testing and training data set? (wtf this cannot be right must check paper).

While the National Phenology Network, and NEON both collect data sets on the timing of life history events, both seem to focus on taxa in dominant physiognomic classes such as trees. These data unfortunately do not contain information that I find useful.

```{r}
weibull_estimates <- read.csv(paste0(here(),'/data/processed/weibull_estimates_phen.csv')) %>% 
  filter(taxon != 'Adenolinum lewisii') %>% 
  mutate(taxon = case_when(
    taxon == 'Ligularia bigelovii' ~ 'Senecio bigelovii',
    taxon == 'Zigadenus elegans' ~ 'Anticlea elegans',
    TRUE ~ taxon
  )) %>% 
  mutate(Estimate = 'Weibull-Subalpine')


weibull_estimates <- read.csv(paste0(here(),'/data/processed/weibull_estimates_mixed_montane.csv')) %>% 
  mutate('Estimate' = 'Weibull-Montane') %>% 
  bind_rows(., weibull_estimates)

groundTruth <- read.csv(paste0(here(),'/data/processed/rmbl_oberved_pheno_CaraDonna2014.csv')) %>% 
  dplyr::select(1:2,7:10) %>% 
  filter(species != 'Eriogonum subalpinum') %>%  # this var. majus of the E. umbellatum complex - E. umbellatum already
  # present this cannot be kept as indepedent whence relegated to synonymy under the BIEN schema
  mutate(species = case_when(
    species == 'Amerosedum lanceolatum' ~ 'Sedum lanceolatum',
    species == 'Ligularia bigelovii' ~ 'Senecio bigelovii', 
    species == 'Dugaldia hoopesii' ~ 'Hymenoxys hoopesii',
    species == 'Erysimum asperum' ~ 'Erysimum capitatum', # this misapplied
    species == 'Amelanchier pumila' ~ 'Amelanchier alnifolia',
    species == 'Ligusticum porterii' ~ 'Ligusticum porteri', 
    species == 'Noccaea montana' ~ 'Noccaea fendleri',
    species == 'Descurainia richardsonii' ~ 'Descurainia incana',
    species == 'Lupinus sericeaus' ~ 'Lupinus sericeus',
    species == 'Sedum rosea' ~ 'Rhodiola integrifolia', # this challenging unable to find how
    # this name was missaplied  to this material
    species == 'Taraxacum officionale' ~ 'Taraxacum officinale',
    species == 'Pedicularis bracteoasa' ~ 'Pedicularis bracteosa', 
    TRUE ~ species
    )) %>% 
  filter(species != 'Mertensia fusiformis') %>% 
  mutate(Duration7412 = round(Duration7412,0))
```

Prepare the observed flowering events onto the estimated flowering events. 
```{r}
weibull_estimates_test <- weibull_estimates %>% 
  filter(taxon %in% groundTruth$species,
         event == 'estimate',
         metric %in% c('tenth','fifty','ninety'))  %>% 
  dplyr::select(!contains('hrb'), -event) %>% 
  group_by(taxon) %>% 
  mutate(est_duration = max(DOY) - min(DOY)) %>% 
  mutate(Type = 'Estimate') %>% 
  mutate(metric = case_when(
    metric == 'tenth' ~ 'First',
    metric == 'fifty' ~ 'Peak',
    metric == 'ninety' ~ 'Last',
  )) 

weibull_estimates_test <- split(weibull_estimates_test, weibull_estimates_test$taxon)
goLong <- function(x){
  res <- bind_rows(x[1:nrow(x),], x[nrow(x),])
  res[nrow(res),1] <- 'Duration' 
  res[nrow(res),2] <- res$est_duration[1]
  res <- res[,c(1:4,6:7)]
  return(res)
}
weibull_estimates_test <- bind_rows(lapply(weibull_estimates_test, goLong)) %>% 
  rename_with(~paste0(.,"_est"), c(-taxon, -metric))

groundTruth <- groundTruth %>% 
  rename_with(~ str_remove(., "_7412|7412")) %>% 
  pivot_longer(!family:species, names_to = 'metric', values_to = 'DOY') %>% 
  mutate(Type = 'Observed', sample_size = 0) %>% 
  rename_with(~paste0(.,"_obs"), c(-species, -metric))

spp_needed <- anti_join(groundTruth, weibull_estimates, by = c('species' = 'taxon')) %>% 
  distinct(species)
rm(spp_needed)
```

```{r Download the missing species and process for comparison, eval = F}
# this code was run iteratively to ensure that I could compare phenology estimates for as many of the ground truthed taxa as possible. These species are more of less quarantined from the spatial species. 

library(sf)

ecoregion_bound <- st_read(here('data/processed/SoRo_ecoregion_bound.shp'), quiet = T)

ecoregions <- read_sf(here("data/raw/us_eco_l4/us_eco_l4_no_st.shp"), quiet = T) %>% 
  dplyr::select(NA_L3CODE, NA_L3NAME) %>% 
  filter(NA_L3NAME == 'Southern Rockies') %>% 
  st_union() %>% 
  st_transform(5070)

spp_needed[9,1] <- 'Lupinus sericeus'
species <- spp_needed %>% pull(species) # Log reg predicted species
ecoregion_bound <- st_as_sf(ecoregion_bound)
ecoregion_bound <- st_transform(ecoregion_bound, 4326)
eco_bound_bbox <- st_bbox(ecoregion_bound)

spp_occurrences <- BIEN::BIEN_occurrence_box(
    min.lat = eco_bound_bbox[2],
    max.lat = eco_bound_bbox[4],
    min.long = eco_bound_bbox[1],
    max.long = eco_bound_bbox[3],
    species = species,
    cultivated = F,
    natives.only = F
)

spp_occurrences1 <- spp_occurrences %>% 
  filter(datasource != c('VegBank', 'FIA')) %>% 
  dplyr::select(scrubbed_species_binomial, latitude, longitude, date_collected) %>% 
  st_as_sf(coords = c(x= "longitude", y = "latitude"), crs = 4326) %>% 
  st_transform(4269)

ecoregion_bound <- st_transform(ecoregion_bound, 4269) # files come in a box, reduce extent to eco-region
spp_occurrences1 <- st_intersection(ecoregion_bound, spp_occurrences1)

fname <- paste0('BIEN_occ_rec_PHEN', Sys.Date(), '.shp')
st_write(spp_occurrences1, here("data/raw", fname), append = F)
rm(spp_occurrences, spp_occurrences1, eco_bound_bbox, fname)
```


- R NOTE: Looks like we want to be able to produce a plot which ALSO contains the lower and upper confidence estimates. 
```{r Prepare Confidence Intervals for plotting}

we <- weibull_estimates %>% 
  filter(taxon %in% groundTruth$species,
         event %in% c('low_ci', 'high_ci'),
         metric %in% c('tenth','fifty','ninety'))  %>% 
  dplyr::select(!contains('hrb')) %>% 
  group_by(taxon) %>% 
  mutate(est_duration = max(DOY) - min(DOY)) %>% 
  mutate(metric = case_when(
    metric == 'tenth' ~ 'First',
    metric == 'fifty' ~ 'Peak',
    metric == 'ninety' ~ 'Last',
  )) %>% 
  dplyr::select(event:taxon) 

low <- we %>% filter(event == 'low_ci') %>% 
  rename('DOY_low' = DOY) %>% 
  dplyr::select(-event)
ci <- we %>% filter(event == 'high_ci') %>% 
  rename('DOY_high' = DOY) %>% 
  dplyr::select(-event) %>% 
  left_join(., low, by = c('taxon', 'metric'))

rm(low, we)
rm(list = lsf.str())
```

# Run Analysis


NOTE WOULD BE NICE TO HAVE A WAY TO QUANTIZE RESIDUALS ALA RMSE - BUT VARIABLES ARE NOT INDEPENDENT ; BOTH ARE PRODUCTS OF THE AXES OF THE SAME UNANALYZED INDEPDENENT VARIABLES. IT APPEARS THERE IS NO METHOD FOR DISPLAYING THE FIT OF CORRELATION TEST/RESULTS... HMMM

```{r Correlation Testing, warning = F}

regression_data <- left_join(groundTruth, weibull_estimates_test, 
                             by = c('species' = 'taxon', 'metric')) %>% 
  left_join(., ci, by = c('species' = 'taxon', 'metric'))

rd_start <- regression_data %>% 
  filter(metric == 'First')
rd_peak <- regression_data %>% 
  filter(metric == 'Peak')
rd_last <- regression_data %>% 
  filter(metric == 'Last')
rd_dur <- regression_data %>% 
  filter(metric == 'Duration')

# START
summary( start_lm <- lm(rd_start$DOY_est ~ rd_start$DOY_obs + rd_start$sample_size_est) )
print( start_lm_rmse <- sqrt(mean((rd_start$DOY_obs - rd_start$DOY_est)^2)) / 
         (max(rd_start$DOY_obs) - min(rd_start$DOY_est)) )
start_kt <- cor.test(rd_start$DOY_est, rd_start$DOY_obs, method="kendall")
summary(start_lm)$adj.r.squared


# PEAK
summary( peak_lm <- lm(rd_peak$DOY_est ~ rd_peak$DOY_obs + rd_start$sample_size_est) )
print( peak_lm_rmse <- sqrt(mean((rd_peak$DOY_obs - rd_peak$DOY_est)^2)) / 
         (max(rd_peak$DOY_obs) - min(rd_peak$DOY_est)) )
peak_kt <- cor.test(rd_peak$DOY_est, rd_peak$DOY_obs, method="kendall")
summary(peak_lm)$adj.r.squared
summary(peak_lm)

# END
summary( last_lm <- lm(rd_last$DOY_est ~ rd_last$DOY_obs + rd_start$sample_size_est) )
print( last_lm_rmse <- sqrt(mean((rd_last$DOY_obs - rd_last$DOY_est)^2)) / 
         (max(rd_last$DOY_obs) - min(rd_last$DOY_est)) )
end_kt <- cor.test(rd_last$DOY_est, rd_last$DOY_obs, method="kendall")
summary(last_lm)$adj.r.squared

# DURATION
summary( duration_lm <- lm(rd_dur$DOY_est ~ rd_dur$DOY_obs + rd_start$sample_size_est ) )
print( dur_lm_rmse <- sqrt(mean((rd_dur$DOY_obs - rd_dur$DOY_est)^2)) / 
         (max(rd_dur$DOY_obs) - min(rd_dur$DOY_est)) )
dur_kt <- cor.test(rd_dur$DOY_est, rd_dur$DOY_obs, method="kendall")
summary(duration_lm)$adj.r.squared

rm(regression_data, weibull_estimates, ci)
```

```{r Plot Correlation, warning = F}


png(file = paste0(here(), '/results/plots/phenology_estimates_10_90.png'), 
    width = 6, height = 6, pointsize = 4, units = 'cm', res = 300)

par(pty="s", mgp=c(2,1,0),  mfrow = c(2,2), mar = c(3,3,2,2))

plotCI(x = rd_start$DOY_obs,  rd_start$DOY_est,
       li = rd_start$DOY_low, ui = rd_start$DOY_high, 
       pch=20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(140,220), ylim = c(140,220), axes = F,
       main = "First Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(148, 214, expression("adj. r"^2*" = " ))
text(162, 214, paste0(round(summary(start_lm)$adj.r.squared, 3)))
text(155, 208, paste0("tau = ", round(start_kt[["estimate"]][["tau"]], 3)))
text(155, 202, "p < 0.001")
abline(start_lm)


plotCI(x = rd_peak$DOY_obs,  rd_peak$DOY_est,
       li = rd_peak$DOY_low, ui = rd_peak$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(160,240), ylim = c(160,240), axes = F,
       main = "Peak Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(168, 234, expression("adj. r"^2*" = " ))
text(182, 234, paste0(round(summary(peak_lm)$adj.r.squared, 3)))
text(175, 228, paste0("tau = ", round(peak_kt[["estimate"]][["tau"]], 3)))
text(175, 222, "p < 0.001")
abline(peak_lm)

plotCI(x = rd_last$DOY_obs,  rd_last$DOY_est,
       li = rd_last$DOY_low, ui = rd_last$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(160,250), ylim = c(160,250), axes = F,
       main = "Last Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(228, 179, expression("adj. r"^2*" = " ))
text(242, 179, paste0(round(summary(last_lm)$adj.r.squared, 3)))
text(235, 174, paste0("tau = ", round(end_kt[["estimate"]][["tau"]], 3)))
text(235, 168, "p < 0.001")
abline(last_lm)

plotCI(x = rd_dur$DOY_obs,  rd_dur$DOY_est,
       li = rd_dur$DOY_low, ui = rd_dur$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(10,90), ylim = c(10,90), axes = F,
       main = "Duration",
       xlab = "Observed (Days)", ylab = "Estimated (Days)")
axis(1)
axis(2)
text(70, 26, expression("adj. r"^2*" = " ))
text(83, 26, paste0(round(summary(duration_lm)$adj.r.squared, 3)))
text(75, 21, paste0("tau = ", round(dur_kt[["estimate"]][["tau"]], 3)))
text(75, 16, paste0("p = ", round(dur_kt[["p.value"]], 3)))
abline(duration_lm)

dev.off()

rm(rd_start, rd_peak, rd_last, rd_dur, groundTruth, weibull_estimates_test,
   duration_lm, end_lm, last_lm, peak_lm, start_lm,
   dur_lm_rmse, last_lm_rmse, peak_lm_rmse, start_lm_rmse,
   dur_kt, end_kt, last_kt, peak_kt, start_kt)
```


CaraDonna, Paul J.; Inouye, David W. (2016): Phenological responses to climate change do not exhibit phylogenetic signal in a subalpine plant community. Wiley. Collection. https://doi.org/10.6084/m9.figshare.c.3307416.v1
