---
title: "evaluate_phenology_predictions"
author: "steppe"
date: '2022-04-23'
output: html_document
---


```{r load libraries, results='hide', message=F, warning=F}
library(here)
library(lubridate)
library(plotrix)
source(here::here('scripts/phen_functions.R'))
library(tidyverse) #data tidying
```

We can only test the accuracy of the Weibull method on a subset of cerca 60 species we have created estimates for. 
An interesting component of the Weibull development paper is that the authors did not actually attempt to estimate phenological events using herbarium records and then test them using the RMBL data set, but rather used the same testing and training data set? 

While the National Phenology Network, and NEON both collect data sets on the timing of life history events, both seem to focus on taxa in dominant physiognomic classes such as trees. These data unfortunately do not contain information that I find useful.

```{r}
weibull_estimates <- read.csv(paste0(here(),'/data/processed/weibull_estimates_phen.csv')) %>% 
  filter(taxon != 'Adenolinum lewisii') %>% 
  mutate(taxon = case_when(
    taxon == 'Ligularia bigelovii' ~ 'Senecio bigelovii',
    taxon == 'Zigadenus elegans' ~ 'Anticlea elegans',
    TRUE ~ taxon
  )) %>% 
  mutate(Estimate = 'Weibull-Subalpine')


weibull_estimates <- read.csv(paste0(here(),'/data/processed/weibull_estimates_mixed_montane.csv')) %>% 
  mutate('Estimate' = 'Weibull-Montane') %>% 
  bind_rows(., weibull_estimates) %>% 
  mutate(taxon = if_else(taxon == 'Arenaria congesta', 'Eremogone congesta', taxon))

groundTruth <- read.csv(paste0(here(),'/data/processed/rmbl_oberved_pheno_CaraDonna2014.csv')) %>% 
  dplyr::select(1:2,7:10) %>% 
  filter(species != 'Eriogonum subalpinum') %>%  # this var. majus of the E. umbellatum complex - E. umbellatum already
  # present this cannot be kept as independent whence relegated to synonymy under the BIEN schema
  mutate(species = case_when(
    species == 'Arenaria congesta' ~ 'Eremogone congesta', 
    species == 'Amerosedum lanceolatum' ~ 'Sedum lanceolatum',
    species == 'Ligularia bigelovii' ~ 'Senecio bigelovii', 
    species == 'Dugaldia hoopesii' ~ 'Hymenoxys hoopesii',
    species == 'Erysimum asperum' ~ 'Erysimum capitatum', # this misapplied
    species == 'Amelanchier pumila' ~ 'Amelanchier alnifolia',
    species == 'Ligusticum porterii' ~ 'Ligusticum porteri', 
    species == 'Noccaea montana' ~ 'Noccaea fendleri',
    species == 'Descurainia richardsonii' ~ 'Descurainia incana',
    species == 'Lupinus sericeaus' ~ 'Lupinus sericeus',
    species == 'Sedum rosea' ~ 'Rhodiola integrifolia', # this challenging unable to find how
    # this name was misapplied  to this material
    species == 'Taraxacum officionale' ~ 'Taraxacum officinale',
    species == 'Pedicularis bracteoasa' ~ 'Pedicularis bracteosa', 
    TRUE ~ species
    )) %>% 
  filter(species != 'Mertensia fusiformis') %>% 
  mutate(Duration7412 = round(Duration7412,0))
```

Prepare the observed flowering events onto the estimated flowering events. 
```{r}
weibull_estimates_test <- weibull_estimates %>% 
  filter(taxon %in% groundTruth$species,
         event == 'estimate',
         metric %in% c('tenth','fifty','ninety'))  %>% 
  dplyr::select(!contains('hrb'), -event) %>% 
  group_by(taxon) %>% 
  mutate(est_duration = max(DOY) - min(DOY)) %>% 
  mutate(Type = 'Estimate') %>% 
  mutate(metric = case_when(
    metric == 'tenth' ~ 'First',
    metric == 'fifty' ~ 'Peak',
    metric == 'ninety' ~ 'Last',
  )) 

weibull_estimates_test <- split(weibull_estimates_test, weibull_estimates_test$taxon)
goLong <- function(x){
  res <- bind_rows(x[1:nrow(x),], x[nrow(x),])
  res[nrow(res),1] <- 'Duration' 
  res[nrow(res),2] <- res$est_duration[1]
  res <- res[,c(1:4,6:7)]
  return(res)
}
weibull_estimates_test <- bind_rows(lapply(weibull_estimates_test, goLong)) %>% 
  rename_with(~paste0(.,"_est"), c(-taxon, -metric))

groundTruth <- groundTruth %>% 
  rename_with(~ str_remove(., "_7412|7412")) %>% 
  pivot_longer(!family:species, names_to = 'metric', values_to = 'DOY') %>% 
  mutate(Type = 'Observed', sample_size = 0) %>% 
  rename_with(~paste0(.,"_obs"), c(-species, -metric))

rm(spp_needed)
```

```{r Download the missing species and process for comparison, eval = F}
# this code was run iteratively to ensure that I could compare phenology estimates for as many of the ground truthed taxa as possible. These species are more of less quarantined from the spatial species. 

library(sf)

ecoregion_bound <- st_read(here('data/processed/SoRo_ecoregion_bound.shp'), quiet = T)

ecoregions <- read_sf(here("data/raw/us_eco_l4/us_eco_l4_no_st.shp"), quiet = T) %>% 
  dplyr::select(NA_L3CODE, NA_L3NAME) %>% 
  filter(NA_L3NAME == 'Southern Rockies') %>% 
  st_union() %>% 
  st_transform(5070)

species <- spp_needed %>% pull(species) # Log reg predicted species
ecoregion_bound <- st_as_sf(ecoregion_bound)
ecoregion_bound <- st_transform(ecoregion_bound, 4326)
eco_bound_bbox <- st_bbox(ecoregion_bound)

spp_occurrences <- BIEN::BIEN_occurrence_box(
    min.lat = eco_bound_bbox[2],
    max.lat = eco_bound_bbox[4],
    min.long = eco_bound_bbox[1],
    max.long = eco_bound_bbox[3],
    species = species,
    cultivated = F,
    natives.only = F
)

spp_occurrences1 <- spp_occurrences %>% 
  filter(datasource != c('VegBank', 'FIA')) %>% 
  dplyr::select(scrubbed_species_binomial, latitude, longitude, date_collected) %>% 
  st_as_sf(coords = c(x= "longitude", y = "latitude"), crs = 4326) %>% 
  st_transform(4269)

ecoregion_bound <- st_transform(ecoregion_bound, 4269) # files come in a box, reduce extent to eco-region
spp_occurrences1 <- st_intersection(ecoregion_bound, spp_occurrences1)

fname <- paste0('BIEN_occ_rec_PHEN', Sys.Date(), '.shp')
st_write(spp_occurrences1, here("data/raw", fname), append = F)
rm(spp_occurrences, spp_occurrences1, eco_bound_bbox, fname)
```

- R NOTE: Looks like we want to be able to produce a plot which ALSO contains the lower and upper confidence estimates. 
```{r Prepare Confidence Intervals for plotting}

we <- weibull_estimates %>% 
  filter(taxon %in% groundTruth$species,
         event %in% c('low_ci', 'high_ci'),
         metric %in% c('tenth','fifty','ninety'))  %>% 
  dplyr::select(!contains('hrb')) %>% 
  group_by(taxon) %>% 
  mutate(est_duration = max(DOY) - min(DOY)) %>% 
  mutate(metric = case_when(
    metric == 'tenth' ~ 'First',
    metric == 'fifty' ~ 'Peak',
    metric == 'ninety' ~ 'Last',
  )) %>% 
  dplyr::select(event:taxon) 

low <- we %>% filter(event == 'low_ci') %>% 
  rename('DOY_low' = DOY) %>% 
  dplyr::select(-event)
ci <- we %>% filter(event == 'high_ci') %>% 
  rename('DOY_high' = DOY) %>% 
  dplyr::select(-event) %>% 
  left_join(., low, by = c('taxon', 'metric'))

rm(low, we)
rm(list = lsf.str())
```

# Run Analysis

all we can do is kendalls tau

```{r Correlation Testing, warning = F}

regression_data <- left_join(groundTruth, weibull_estimates_test, 
                             by = c('species' = 'taxon', 'metric')) %>% 
  left_join(., ci, by = c('species' = 'taxon', 'metric'))

rd_start <- regression_data %>% 
  filter(metric == 'First')
rd_peak <- regression_data %>% 
  filter(metric == 'Peak')
rd_last <- regression_data %>% 
  filter(metric == 'Last')
rd_dur <- regression_data %>% 
  filter(metric == 'Duration')

# START
summary( start_lm <- lm(rd_start$DOY_est ~ rd_start$DOY_obs) )
print( start_lm_rmse <- sqrt(mean((rd_start$DOY_obs - rd_start$DOY_est)^2)) / 
         (max(rd_start$DOY_obs) - min(rd_start$DOY_est)) )
start_kt <- cor.test(rd_start$DOY_est, rd_start$DOY_obs, method="kendall")

# PEAK
summary( peak_lm <- lm(rd_peak$DOY_est ~ rd_peak$DOY_obs) )
print( peak_lm_rmse <- sqrt(mean((rd_peak$DOY_obs - rd_peak$DOY_est)^2)) / 
         (max(rd_peak$DOY_obs) - min(rd_peak$DOY_est)) )
peak_kt <- cor.test(rd_peak$DOY_est, rd_peak$DOY_obs, method="kendall")

# END
summary( last_lm <- lm(rd_last$DOY_est ~ rd_last$DOY_obs) )
print( last_lm_rmse <- sqrt(mean((rd_last$DOY_obs - rd_last$DOY_est)^2)) / 
         (max(rd_last$DOY_obs) - min(rd_last$DOY_est)) )
end_kt <- cor.test(rd_last$DOY_est, rd_last$DOY_obs, method="kendall")

# DURATION
summary( duration_lm <- lm(rd_dur$DOY_est ~ rd_dur$DOY_obs ) )
print( dur_lm_rmse <- sqrt(mean((rd_dur$DOY_obs - rd_dur$DOY_est)^2)) / 
         (max(rd_dur$DOY_obs) - min(rd_dur$DOY_est)) )
dur_kt <- cor.test(rd_dur$DOY_est, rd_dur$DOY_obs, method="kendall")

regression_data %>% 
  distinct(species, .keep_all = T) %>%
  summarize(med = median(sample_size_est),
            mean = round(mean(sample_size_est), 1))

rm(regression_data, weibull_estimates)
```


```{r Repeat process to compare first last and mid point of flowering }

jeo_groundtruth <- read.csv(paste0(here(),'/data/processed/Bombus_flower_ranks_2015.csv')) %>% 
  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank) %>% 
  group_by(plant.species) %>% 
  arrange(doy) %>% 
  slice(1, n()) %>% 
  select(species = plant.species,  DOY_obs = doy) %>% 
  mutate(
    metric = rep(c('First', 'Last'), times = 1),
    species = str_replace_all(species, '[.]', " "),
    species = case_when(
      species == 'Adenolinum lewisii' ~ 'Linum lewisii', 
      species == 'Amerosedum lanceolatum' ~ 'Sedum lanceolatum',
      species == 'Ligularia bigelovii' ~ 'Senecio bigelovii', 
      species == 'Dugaldia hoopesii' ~ 'Hymenoxys hoopesii',
      species == 'Erysimum asperum' ~ 'Erysimum capitatum', # this misapplied
      species == 'Amelanchier pumila' ~ 'Amelanchier alnifolia',
      species == 'Ligusticum porterii' ~ 'Ligusticum porteri', 
      species == 'Noccaea montana' ~ 'Noccaea fendleri',
      species == 'Descurainia richardsonii' ~ 'Descurainia incana',
      species == 'Lupinus sericeaus' ~ 'Lupinus sericeus',
      species == 'Taraxacum officionale' ~ 'Taraxacum officinale',
      species == 'Pedicularis bracteoasa' ~ 'Pedicularis bracteosa', 
      TRUE ~ species
    )) %>% 
  filter(! species %in% c('Salix sp', 'Carex sp', 'Epilobium sp', 'Eriogonum subalpinum'))

jeo_groundtruth <- jeo_groundtruth %>% 
  inner_join(., select( weibull_estimates_test, DOY_est, metric, sample_size_est) ,
            by = c('species' = 'taxon', 'metric')) %>% 
  left_join(., ci, by = c('species' = 'taxon', 'metric'))

jeo_start <- filter(jeo_groundtruth, metric == 'First')
jeo_last <- filter(jeo_groundtruth, metric == 'Last')

# How many of the species are repeated between the JEO dataset and long term dataset ?
lt_species <-  distinct(groundTruth, species) %>% 
           pull()

jeo_start %>% # 47 of the 47 species are repeated... 
  filter(species %in% lt_species)

summary( start_lm_jeo <- lm(jeo_start$DOY_est ~ jeo_start$DOY_obs) )
print( start_lm_rmse_jeo <- sqrt(mean((jeo_start$DOY_obs - jeo_start$DOY_est)^2)) / 
         (max(jeo_start$DOY_obs) - min(jeo_start$DOY_est)) )
start_kt_jeo <- cor.test(jeo_start$DOY_est, jeo_start$DOY_obs, method="kendall")

summary( last_lm_jeo <- lm(jeo_last$DOY_est ~ jeo_last$DOY_obs) )
print( last_lm_rmse_jeo <- sqrt(mean((jeo_last$DOY_obs - jeo_last$DOY_est)^2)) / 
         (max(jeo_last$DOY_obs) - min(jeo_last$DOY_est)) )
end_kt_jeo <- cor.test(jeo_last$DOY_est, jeo_last$DOY_obs, method="kendall")

# The survey finished before field season did. we will predict the values for the 
# trend line onto a new dataframe for plotting

rm(ci)
```


```{r Plot Correlation, warning = F}

png(file = paste0(here(), '/results/plots/phenology_estimates_10_90.png'), 
    width = 6, height = 6, pointsize = 4, units = 'cm', res = 300)

par(pty="s", mgp=c(2,1,0),  mfrow = c(2,2), mar = c(3,3,2,2))

plotCI(x = rd_start$DOY_obs,  rd_start$DOY_est,
       li = rd_start$DOY_low, ui = rd_start$DOY_high, 
       pch=20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(140,220), ylim = c(140,220), axes = F,
       main = "First Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(155, 208, paste0("tau = ", round(start_kt[["estimate"]][["tau"]], 3)))
text(155, 202, "p < 0.001")
abline(start_lm)


plotCI(x = rd_peak$DOY_obs,  rd_peak$DOY_est,
       li = rd_peak$DOY_low, ui = rd_peak$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(160,240), ylim = c(160,240), axes = F,
       main = "Peak Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(175, 228, paste0("tau = ", round(peak_kt[["estimate"]][["tau"]], 3)))
text(175, 222, "p < 0.001")
abline(peak_lm)

plotCI(x = rd_last$DOY_obs,  rd_last$DOY_est,
       li = rd_last$DOY_low, ui = rd_last$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(160,250), ylim = c(160,250), axes = F,
       main = "Last Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(235, 174, paste0("tau = ", round(end_kt[["estimate"]][["tau"]], 3)))
text(235, 168, "p < 0.001")
abline(last_lm)

plotCI(x = rd_dur$DOY_obs,  rd_dur$DOY_est,
       li = rd_dur$DOY_low, ui = rd_dur$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(10,90), ylim = c(10,90), axes = F,
       main = "Duration",
       xlab = "Observed (Days)", ylab = "Estimated (Days)")
axis(1)
axis(2)
text(75, 21, paste0("tau = ", round(dur_kt[["estimate"]][["tau"]], 3)))
text(75, 16, paste0("p = ", round(dur_kt[["p.value"]], 3)))
abline(duration_lm)
dev.off()


# now repeat for the jane data

png(file = paste0(here(), '/results/plots/phenology_estimates-jeo_10_90.png'), 
    width = 6, height = 3, pointsize = 4, units = 'cm', res = 300)
par(pty="s", mgp=c(2,1,0),  mfrow = c(1,2), mar = c(3,3,2,2))
plotCI(x = jeo_start$DOY_obs,  jeo_start$DOY_est,
       li = jeo_start$DOY_low, ui = jeo_start$DOY_high, 
       pch=20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(140,220), ylim = c(140,220), axes = F,
       main = "First Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)") 
axis(1)
axis(2)
text(155, 208, paste0("tau = ", round(start_kt_jeo[["estimate"]][["tau"]], 3)))
text(155, 202, "p < 0.001")
abline(start_lm_jeo)

plotCI(x = jeo_last$DOY_obs,  jeo_last$DOY_est,
       li = jeo_last$DOY_low, ui = jeo_last$DOY_high,
       pch = 20, slty = 3, gap = 0.025, scol = 'grey70', sfrac = 0.0, col = 'grey15',
       xlim = c(160,250), ylim = c(160,250), axes = F,
       main = "Last Flower Date",
       xlab = "Observed (DOY)", ylab = "Estimated (DOY)")
axis(1)
axis(2)
text(235, 174, paste0("tau = ", round(end_kt_jeo[["estimate"]][["tau"]], 3)))
text(235, 168, "p < 0.001")
abline(last_lm_jeo)
dev.off()


rm(rd_start, rd_peak, rd_last, rd_dur, groundTruth, weibull_estimates_test,
   duration_lm, end_lm, last_lm, peak_lm, start_lm, end_kt_jeo, jeo_last, jeo_start,
   dur_lm_rmse, last_lm_rmse, peak_lm_rmse, start_lm_rmse, last_lm_jeo, start_kt_jeo, 
   dur_kt, end_kt, last_kt, peak_kt, start_kt, start_lm_jeo, jeo_groundtruth,
   last_lm_rmse_jeo, start_lm_rmse_jeo
)
```


CaraDonna, Paul J.; Inouye, David W. (2016): Phenological responses to climate change do not exhibit phylogenetic signal in a subalpine plant community. Wiley. Collection. https://doi.org/10.6084/m9.figshare.c.3307416.v1
