---
title: "Extract SDM's to RMBL"
author: "steppe"
date: "3/22/2022"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    includes:
      in_header: float_free.tex
  word_document: default

---

\tableofcontents 
\listoffigures
\listoftables

Finally, we can extract SDMs to the field sites near the Rocky Mountain Biological Laboratory. This is the final step in this portion of the project.
```{r load libraries, results='hide', message=F, warning=F}
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(raster) # raster data
library(here)
library(spdep)
set.seed(12)
```

```{r Create Raster Stack of Predictons, echo = F}
predictions_stack <- paste0(here(), '/results/maps/',
list.files(paste0(here(), '/results/maps/'),
           pattern = "*.tif$" ))
predictions_stack <- raster::stack(predictions_stack)

```

```{r Import Site data, warning=F, echo = F}
sites <- paste0(here(), '/data/processed/', 
list.files(paste0(here(), '/data/processed/'), 
           pattern = 'Bombus_site_occurences.shp'))
sites <- st_read(sites, quiet = T) %>% dplyr::select(site)
sites_cent <- st_centroid(sites)
```

Given that the points more or less follow the lower slopes of the valleys, we will use them to quickly create a elongated polygon feature which occupies the valley. To do this I utilize a Minimum Spanning Tree algorithm, and manually select the neighbors between all sites, and the origin point for the spannning tree.
```{r Create Minimum Spanning Tree to Connect Sites, warning=F, echo = F}
dist_nb <- knn2nb(knearneigh(sites_cent, k=3))
listw <- nb2listw(dist_nb, style="B")
mst.bh <- mstree(listw,2)

#plot(st_geometry(sites), border=gray(.5))
#plot(mst.bh, coordinates(as(sites, "Spatial")), col=2, 
#     cex.lab=.6, cex.circles=0.035, fg="blue", add=TRUE)

mst.bh <- mst.bh[,1:2]
colnames(mst.bh) <- c('From','To')

coords <- sites_cent %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>% 
  st_drop_geometry()

mst.bh <- as.data.frame(mst.bh)
lines <- data.frame(matrix(ncol=1, nrow=5))
for (i in 1:nrow(mst.bh)){
   lines[i,] <-  st_linestring(matrix(c(coords[mst.bh[i,1],2], coords[mst.bh[i,2],2], 
                       coords[mst.bh[i,1],3], coords[mst.bh[i,2],3]),
                     2, 2)) %>% 
     st_sfc() %>% 
     st_as_sf(crs = 32613)
}

lines <- lines %>% st_as_sf(crs = 32613)
st_crs(lines) <- 32613
st_geometry(lines) <- 'geometry' 

rm(dist_nb, coords, listw, mst.bh, i, sites_cent)
``` 

```{r buffer minimum spanning tree, echo = F, message = F, warning = F}
study_area <- st_buffer(lines, dist = 400, endCapStyle = 'ROUND', 
                        joinStyle = 'ROUND') %>% 
  st_union() %>% 
  st_as_sf()

site_bbox <- study_area %>% 
  st_buffer(250) %>% 
  st_transform(4326) %>% 
  st_bbox(sites)
names(site_bbox) <- c('left', 'bottom', 'right', 'top')
basemap <- ggmap::get_stamenmap(site_bbox, zoom = 14, maptype = "terrain", messaging = F)

ggmap::ggmap(basemap) +
  geom_sf(data = study_area,
          alpha = 0.5,
          fill = "chocolate2",
          color = "chocolate2",
          show.legend = "line",
          size = 1.2,
          inherit.aes = FALSE) +
  coord_sf(crs = st_crs(4326)) 

rm(lines)
```

```{r, eval = F}

j_sites_buf <- study_area %>% 
  st_transform(32613) %>% 
  st_buffer(1000)

sdp_prods <- 
  read.csv("https://www.rmbl.org/wp-content/uploads/2021/04/SDP_product_table_4_26_2021.csv")

slope_angle_uri <- sdp_prods %>%
   filter(Product == 'Topographic Slope Angle') %>%
   pull(Data.URL)
slope_angle_path <- paste0("/vsicurl/", slope_angle_uri)
slope_angle <- raster(slope_angle_path, progress='text')
slope_angle <- crop(slope_angle, extent(j_sites_buf))

writeRaster(slope_angle, paste0(here(),"/data/raw/Slope_angle"))
rm(slope_angle_uri, slope_angle_path, slope_angle)
```


```{r Extract SDM predictions to polygons underlaying the area, echo = F}
study_area <- st_transform(study_area, st_crs(predictions_stack[[1]]))
extracted_values <- terra::extract(predictions_stack, study_area)
extracted_values <- data.frame(extracted_values)
names(extracted_values) <- gsub("_glm.*", "", names(extracted_values))
names(extracted_values) <- make.unique(names(extracted_values), "_")
sample_size <- nrow(extracted_values) 

extracted_values <- extracted_values %>% summarise(across(1:ncol(.), summary))
rownames(extracted_values) <- names(summary(c(1))) 
extracted_values <- extracted_values %>% 
  rownames_to_column('Stat') %>%  
  pivot_longer(!Stat) %>% 
  mutate(value = as.numeric(value))

extracted_values  %>% 
  mutate(value = if_else(value < 0, 0, value)) %>% 
  mutate(value = if_else(value > 1, 1, value)) %>% 
  ggplot() +
  geom_density(aes(value, fill = Stat), position="stack") +
  labs(title="Summary of Suitable Habitat from SDM's in East River Valley",
        x ="Probability of Suitable Habitat", y = "Density") +
  theme_bw() +
  xlim(0,1)

```


```{r species to remove, echo = F}

removals <- extracted_values %>% 
  filter( Stat == 'Mean' & value <= 0.5) %>% 
  filter(str_detect(name, '_1$',negate= T))

targets <- extracted_values %>% 
  filter(Stat == 'Mean' & value >= 0.5) %>% 
  filter(str_detect(name, '_1$',negate= T))
```


# THIS NEEDS TO REFLECT THAT AN EXCESS OF SPECIES WERE MODELLED AT RMBL, AS WE STARTED WITH THE VASCULAR FLORA, BEFORE GOING BACK TO TRY THE NEW APPROACH. FOR THE ANALYSIS SPECIES WHICH WERE NOT RECOVERED VIA THE DISTANCE SEARCH AND REGRESSION NEED BE REMOVED.

```{r write out target taxa, echo = F}

targets1 <- targets %>% 
  mutate(approach = if_else(str_detect(name, '_ml_2022'), 'ML', 'LM')) %>% 
  mutate(name = str_remove(name, '_ml_2022.*$'))

removals1 <- removals %>% 
  mutate(approach = if_else(str_detect(name, '_ml_2022'), 'ML', 'LM')) %>% 
  mutate(name = str_remove(name, '_ml_2022.*$'))

consensus <- targets %>% 
  group_by(name) %>% 
  filter(n() == 2) %>% 
  distinct(name) %>% 
  mutate(Algo = 'consensus')

target <- targets %>% 
  group_by(name) %>% 
  filter(n() == 1) %>% 
  mutate(Algo = approach) %>% 
  bind_rows(., consensus) %>% 
  dplyr::select(-approach, -Stat, -value) %>% 
  arrange(name)

write.csv(target, paste0(here(), '/data/processed/taxa_predicted_rmbl.csv'), row.names = F)
write.csv(removals1, paste0(here(), '/data/processed/taxa_not_predicted_rmbl.csv'), row.names = F)
write.csv(targets1, paste0(here(), '/data/processed/taxa_predicted_rmbl_stats_only.csv'), row.names = F)
```



















```{r}
# determine the true absences and presences for RMBL ONLY.

rmbl_spp <- read.csv(paste0(here(), '/data/processed/rmbl_wfo_synonyms.csv'))# %>% 
  mutate(Node.1 = replace_na(Node.1, 'Vascular')) %>% 
  mutate('PrtyFlrs' = case_when(
    Group != 'angiosperms' ~ 0,
    Group == 'angiosperms' & Order == 'Poales' ~ 0,
    TRUE ~ 1
  )) %>% 
  filter(PrtyFlrs == 1) %>% 
  mutate(name = str_replace_all(Species, " ", "_"))

target <- read.csv(paste0(here(), '/data/processed/taxa_predicted_rmbl.csv')) %>% 
  mutate(name = str_replace(name, 'uva.ursi', "uva_ursi")) %>% 
  mutate(Presence = if_else(name %in% rmbl_spp$name, 1, 0) )


LM_res <- target %>% filter(Algo %in% c('LM', 'consensus')) %>% nrow()
ML_res <- target %>% filter(Algo %in% c('ML', 'consensus')) %>% nrow()

```



```{r}
library(WorldFlora) # maybe you need this
WFO.remember()

rmbl_spp <- filter(target, !is.na(name)) %>% pull(name)
rmbl_fam <- wfo_family_finder(rmbl_spp)
target_fam <- wfo_family_finder(target_spp)


??wfo_family_finder
```


REFERENCES:

R. C. Prim (1957) Shortest connection networks and some generalisations. In: Bell System Technical Journal, 36, pp. 1389-1401



