---
title: "Extract SDM's to RMBL"
author: "steppe"
date: "3/22/2022"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    includes:
      in_header: float_free.tex
  word_document: default

---

\tableofcontents 
\listoffigures
\listoftables

Finally, we can extract SDMs to the field sites near the Rocky Mountain Biological Laboratory. This is the final step in this portion of the project.
```{r load libraries, results='hide', message=F, warning=F}
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(raster) # raster data
library(here)
library(spdep)
set.seed(12)
```

```{r Create Raster Stack of Predictons, echo = F}
predictions_stack <- paste0(here(), '/results/maps/',
list.files(paste0(here(), '/results/maps/'),
           pattern = "*.tif$" ))
predictions_stack <- raster::stack(predictions_stack)
```

```{r Import Site data, warning=F, echo = F}
sites <- paste0(here(), '/data/processed/', 
list.files(paste0(here(), '/data/processed/'), 
           pattern = 'Bombus_site_occurences.shp'))
sites <- st_read(sites, quiet = T) %>% dplyr::select(site)
sites_cent <- st_centroid(sites)
```

Given that the points more or less follow the lower slopes of the valleys, we will use them to quickly create a elongated polygon feature which occupies the valley. To do this I utilize a Minimum Spanning Tree algorithm, and manually select the neighbors between all sites, and the origin point for the spannning tree.
```{r Create Minimum Spanning Tree to Connect Sites, warning=F, echo = F}
dist_nb <- knn2nb(knearneigh(sites_cent, k=3))
listw <- nb2listw(dist_nb, style="B")
mst.bh <- mstree(listw,2)

#plot(st_geometry(sites), border=gray(.5))
#plot(mst.bh, coordinates(as(sites, "Spatial")), col=2, 
#     cex.lab=.6, cex.circles=0.035, fg="blue", add=TRUE)

mst.bh <- mst.bh[,1:2]
colnames(mst.bh) <- c('From','To')

coords <- sites_cent %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>% 
  st_drop_geometry()

mst.bh <- as.data.frame(mst.bh)
lines <- data.frame(matrix(ncol=1, nrow=5))
for (i in 1:nrow(mst.bh)){
   lines[i,] <-  st_linestring(matrix(c(coords[mst.bh[i,1],2], coords[mst.bh[i,2],2], 
                       coords[mst.bh[i,1],3], coords[mst.bh[i,2],3]),
                     2, 2)) %>% 
     st_sfc() %>% 
     st_as_sf(crs = 32613)
}

lines <- lines %>% st_as_sf(crs = 32613)
st_crs(lines) <- 32613
st_geometry(lines) <- 'geometry' 

rm(dist_nb, coords, listw, mst.bh, i, sites_cent)
``` 

```{r buffer minimum spanning tree, echo = F, message = F, warning = F}
study_area <- st_buffer(lines, dist = 400, endCapStyle = 'ROUND', 
                        joinStyle = 'ROUND') %>% 
  st_union() %>% 
  st_as_sf()

site_bbox <- study_area %>% 
  st_buffer(250) %>% 
  st_transform(4326) %>% 
  st_bbox(sites)
names(site_bbox) <- c('left', 'bottom', 'right', 'top')
basemap <- ggmap::get_stamenmap(site_bbox, zoom = 14, maptype = "terrain", messaging = F)

ggmap::ggmap(basemap) +
  geom_sf(data = study_area,
          alpha = 0.5,
          fill = "chocolate2",
          color = "chocolate2",
          show.legend = "line",
          size = 1.2,
          inherit.aes = FALSE) +
  coord_sf(crs = st_crs(4326)) 

rm(lines)
```

```{r, eval = F}

j_sites_buf <- study_area %>% 
  st_transform(32613) %>% 
  st_buffer(1000)

sdp_prods <- 
  read.csv("https://www.rmbl.org/wp-content/uploads/2021/04/SDP_product_table_4_26_2021.csv")

slope_angle_uri <- sdp_prods %>%
   filter(Product == 'Topographic Slope Angle') %>%
   pull(Data.URL)
slope_angle_path <- paste0("/vsicurl/", slope_angle_uri)
slope_angle <- raster(slope_angle_path, progress='text')
slope_angle <- crop(slope_angle, extent(j_sites_buf))

writeRaster(slope_angle, paste0(here(),"/data/raw/Slope_angle"))
rm(slope_angle_uri, slope_angle_path, slope_angle)
```


```{r Extract SDM predictions to polygons underlaying the area, echo = F}
study_area <- st_transform(study_area, st_crs(predictions_stack[[1]]))
extracted_values <- terra::extract(predictions_stack, study_area)
extracted_values <- data.frame(extracted_values)
names(extracted_values) <- gsub("_glm.*", "", names(extracted_values))
names(extracted_values) <- make.unique(names(extracted_values), "_")
sample_size <- nrow(extracted_values) 

extracted_values <- extracted_values %>% summarise(across(1:ncol(.), summary))
rownames(extracted_values) <- names(summary(c(1))) 
extracted_values <- extracted_values %>% 
  rownames_to_column('Stat') %>%  
  pivot_longer(!Stat) %>% 
  mutate(value = as.numeric(value))

extracted_values  %>% 
  mutate(value = if_else(value < 0, 0, value)) %>% 
  mutate(value = if_else(value > 1, 1, value)) %>% 
  ggplot() +
  geom_density(aes(value, fill = Stat), position="stack") +
  labs(title="Summary of Suitable Habitat from SDM's in East River Valley",
        x ="Probability of Suitable Habitat", y = "Density") +
  theme_bw() +
  xlim(0,1)

```


```{r species to remove, echo = F}

removals <- extracted_values %>% 
  filter( Stat == 'Mean' & value <= 0.5) %>% 
  filter(str_detect(name, '_1',negate= T))

targets <- extracted_values %>% 
  filter(Stat == 'Mean' & value >= 0.5) %>% 
  filter(str_detect(name, '_1',negate= T))
```

```{r write out target taxa, echo = F}

targets <- targets %>% 
  mutate(approach = if_else(str_detect(name, '_ml_2022'), 'ML', 'LM')) %>% 
  mutate(name = str_remove(name, '_ml_2022.*$'))

consensus <- targets %>% 
  group_by(name) %>% 
  filter(n() == 2) %>% 
  distinct(name) %>% 
  mutate(Algo = 'consensus')

target <- targets %>% 
  group_by(name) %>% 
  filter(n() == 1) %>% 
  mutate(Algo = approach) %>% 
  bind_rows(., consensus) %>% 
  dplyr::select(-approach, -Stat, -value) %>% 
  arrange(name)

write.csv(target, paste0(here(), '/data/processed/taxa_predicted_rmbl.csv'), row.names = F)

```


REFERENCES:

R. C. Prim (1957) Shortest connection networks and some generalisations. In: Bell System Technical Journal, 36, pp. 1389-1401



