---
title: "Generating_psuedo_absences"
author: "steppe"
date: "3/5/2022"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    includes:
      in_header: float_free.tex
  word_document: default

---

\tableofcontents 
\listoffigures
\listoftables


```{r load libraries, results='hide', message=F, warning=F}
library(sdm) 
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(raster) # raster data
library(here)

source(here::here('scripts/sdm_functions.R'))
set.seed(12)
```


```{r}
spp_occurrences_reg_random <- st_read(here("data/processed/spp_occurences_regression.shp"), quiet = T)
spp_occurrences <- st_read(here("data/processed/spp_occurences.shp"), quiet = T)
ecoregion_bound <- st_read(here('data/processed/SoRo_ecoregion_bound.shp'), quiet = T)
BLM_CP <- st_read(here("data/raw/blm_ownership/BLM_ownership_in_study_area.shp"), quiet = T)
BLM_AIM <- st_read(here("data/raw/AIM_plants_Colorado_Basin.shp"), quiet = T) %>% 
  dplyr::select(PlotKey, binomil) %>% 
  rename("binomial" = binomil) %>% 
  group_by(PlotKey)
```

```{r Prepare spatial products for generating Psuedo-absence for Regression via geography}

ecoregion_bound <- st_transform(ecoregion_bound, 5070)
blm_cp_s <- BLM_CP %>% 
  st_transform(5070) %>% 
  st_buffer(400) # we will buffer the BLM fine borders to 400 m as these areas are virtually always in the same landscape feature the method was designed to capture.

blm_cp_s <- rmapshaper::ms_simplify(blm_cp_s, keep = 0.05) %>% # we will simplify the shapes of the BLM land. 
  st_make_valid()

st_erase = function(x, y) st_difference(x, st_union(st_combine(y))) # function to erase the BLM land from area. 
blm_cp_s <- st_erase(ecoregion_bound, blm_cp_s) %>% 
  st_cast(to = "POLYGON") # turn all polygons in the feature to their own vector objects

blm_cp_s$AREA <- st_area(blm_cp_s) # calculate the area of polygons to remove minor inclusions (e.g. private land on post-stamped BLM)
blm_cp_s <- blm_cp_s %>% 
  mutate(AREA = as.numeric(AREA)) %>% # units in meters squared
  filter(AREA >= 1000000) %>%  # > 1,000 hectare. 
  st_union()
```

```{r load and prep raster stacks, echo = F}
names_vec <- c('PCL_MAN','PCL_SE1','PCL_SE2','PCL_SE3','PCM_BFP','PCM_CMD',
                  'PCM_DD5', 'PCM_MAP','PCM_PAS','PCM_TD','PCT_ECO','PCT_EFW',
                  'PLC_HRB', 'PLC_TRE', 'PLC_URB', 'PSL_BDR', 'PSL_CAR', 
                  'PSL_PHO', 'PSL_SND','PSL_TUS','PSL_TWL','PTP_ELV','PTP_RLV',
                  'PTP_SLP','PTP_WET','PTP_ASP','PTP_SOL', 'PCL_MRA','PSW_DIS',
                  'PSW_OCC','PCO_XSC', 'PCO_YSC')

WPDPV <- raster::stack("/hdd/Geospatial_data/Western_Plant_Predictors/SoRo_wppr_vers_two.tif")
names(WPDPV) <- names_vec
target <- c(2, 4,9, 10, 13, 14, 16, 19, 20, 23, 24, 25, 26, 27, 29, 30)
target_ml <- c(1:10, 13:14, 16:20, 22:30)

WPDPV2 <- WPDPV[[(target)]]
WPDPV1 <- WPDPV[[(target_ml)]]

rm(WPDPV, target, target_ml, names_vec)
```

# Table of Contents

Methods will be broadly be defined as four main components. 

1) Data Cleaning, and Processing
  + Part 1: Define the study extent.
  + Part 2: Collect, and clip input Independent variables to this extent. 
  + Part 3: Remove predictor variables which exhibit collinearity for the linear regressions. 
  + Part 4: Assemble Presence Records for the extent, formatting BLM data
2) Tests for Spatial Auto-Correlation
  + Part 1: Prepare data, including creating grids, and generated neighbor objects
  + Part 2: Calculate Moran's I and Geary's C
  + Part 3: Sub-sample the presence records to reduce Spatial auto-correlation
3 **Generating Pseudo - absences**
  + Part 1: Append true absence records to the data frames for each Species which will undergo ML and Linear Regression
  + Part 2: Ensure each Species data frame which will be subjected to LR has 1,000 records, boosted by pseudo-absences.
  + Part 3: Ensure each species data frame which will be subjected to ML has no. Presence == no. Pseudo-absence records factoring in environmental dissimilarity.

# 3: Pseudo-Absences

Many of these models will require pseudo-absences for the generation of predictions. We will use both some true-absences, and pseudo-absences to generate the number of records required to satisfy the models. 

## 3.1:  Append true absence records to the data frames for each Species which will undergo ML and Linear Regression

We will load these records with essentially true absences. While we cannot be sure a 400m^2 cell is lacking a record, this is about as good as we can get. 

```{r true absences Regression part 1, eval = F}
prop_blm <- as.numeric((st_area(BLM_CP))/(st_area(ecoregion_bound)))
spp_occurrences <- spp_occurrences %>% st_transform(32613)

AIM_points <- BLM_AIM %>% 
  dplyr::select(-binomial) %>% 
  distinct(PlotKey, .keep_all = T) %>% 
  mutate(occurrence = 0)
AIM_points <- AIM_points %>% 
  st_transform(32613)

spp_occurrence_list <- split(as.data.frame(spp_occurrences), f = as.data.frame(spp_occurrences$binomial))
PA_REG_DF <- lapply(spp_occurrence_list, true_absence_REG)
```

```{r True absences Regression, eval = F}
spp_occurrence_REG_list <- split(spp_occurrences_reg_random,  ~ binomial)
PA_REG_DF <- lapply(spp_occurrence_REG_list, true_absence_REG)
```

When ML seems to perform better than traditional techniques, I still think it is worth building an ensemble forecast from both GLM + GAM, and RF + BRTs, saving them and having a user evaluate them. Both of these sets will need different replicates, and different pseudo-absences. 

## 3.2 Ensure each Species data frame which will be subjected to Linear regression has 1,000 records, boosted by pseudo-absences.

In SDMs built from GLM's and GAM's a much higher number of (pseudo)-absences are required, 10,000 has been suggesed as a good rule of thumb, but this lead to results where we could not generalize occurrence from the training data. We will cut this rule an order of magnitude to 1000 points. When selecting these absences, we will allow for the placement to be randomly distributed across the geographic extent of analysis (excluding BLM land - remember these areas are accounted for by AIM). We will use minimum offset distances to avoid sampling the same raster cell twice. So we do not actually spatially explicitly sample here, we will just sample the sub-setted data frame from the raster 1,000 times, after removing the presences. 

```{r Psuedo-absence for Regression via geography, eval = F}

PA_REG_DF <- PA_REG_DF %>% 
  map(~st_buffer(.x, 10000)) %>% 
  map(~st_transform(.x, 5070))

Sys.time()
test <- lapply(PA_REG_DF, random_PA_spatial) # about 30 minutes
Sys.time()
test1 <- data.table::rbindlist(test) %>%  st_as_sf()

st_write(test1, here("data/processed/Pseudo_asbences_regression_geographically_generated.shp"), append = F)
```

```{r, echo = F, warning = F, message = F}
rm(test, test1)
```

Now to extract the climate variables for each record from the Pseudo-Absences
```{r, Extract values for climate variables, echo = F, eval = F}
PA_REG_DF <- read_sf(here("data/processed/Pseudo_absences_regression_geographically_generated.shp") )
PA_REG_DF <- PA_REG_DF %>% filter(!st_is_empty(.))
PA_REG_DF1 <- raster::extract(WPDPV2, PA_REG_DF) # start 12:49. 

PA_REG_DF <- cbind(PA_REG_DF, PA_REG_DF1)
PA_REG_DF1 <- PA_REG_DF %>% drop_na(PCL_SE1) 

st_write(PA_REG_DF1,  here("data/processed/Pseudo_absences_regression.shp"), append = F, quiet = T)
rm(PA_REG_DF1)
```

We have these data saved so we do not need to rerun these extractions. 
```{r, Bring in Regression Pseudo-Absences, eval = F}
PA_REG_DF <- st_read(here("data/processed/Pseudo_absences_regression.shp"), quiet = T) # 461647
PA_REG_DF <- PA_REG_DF %>% drop_na(PCL_SE1) 

PA_REG_DF_LIST <- split(PA_REG_DF, ~ binomial)
```

We now have 1,000 Absence/Pseudo-Absence records for each species we want to analyze. As mentioned we initially ran 10k records, but preliminary modeling indicated that this many records swamped the ability of the models to generalize, hence overly conservative predictions. It is plausible that this ratio of absences to presences is swamping the link functions in the regressions (I. Breckheimer, Pers. Comm. May. 2021) 

## C. Part 3: Pseudo-Absence generation for RF & BRT

```{r true absences ML part 1, eval = F}
prop_blm <- as.numeric((st_area(BLM_CP))/(st_area(ecoregion_bound)))
spp_occurrences <- spp_occurrences %>% st_transform(32613)

AIM_points <- BLM_AIM %>% 
  dplyr::select(-binomial) %>% 
  distinct(PlotKey, .keep_all = T) %>% 
  mutate(occurrence = 0)
AIM_points <- AIM_points %>% 
  st_transform(32613)

spp_occurrence_list <- split(as.data.frame(spp_occurrences), f = as.data.frame(spp_occurrences$binomial))
PA_ML_DF <- lapply(spp_occurrence_list, true_absence_ML)
```

```{r, echo = F}
rm(prop_blm, AIM_points, spp_occurrence_list)
```

```{r Psuedo-absence for ML via geography, echo = F, eval = F}

PA_ML_DF <- PA_ML_DF %>% 
  map(~st_as_sf(.x)) %>% 
  map(~st_buffer(.x, 10000)) %>% 
  map(~st_transform(.x, 5070))

Sys.time()
test <- lapply(PA_ML_DF, random_PAML_spatial) # 11 minutes
Sys.time()
test1 <- data.table::rbindlist(test) %>% st_as_sf()

st_write(test1, here("data/processed/Pseudo_absences_ML_geographically_generated.shp"), append = F, quiet = T)
```

Now to extract the climate variables for each record from the Pseudo-Absences
```{r, Extract values for climate variables, echo = F, eval = F}
PA_ML_DF <- read_sf(here("data/processed/Pseudo_absences_ML_geographically_generated.shp") )
PA_ML_DF <- PA_ML_DF %>% 
  filter(!st_is_empty(.)) %>% 
  st_transform(st_crs(WPDPV1))
PA_ML_DF1 <- raster::extract(WPDPV1, PA_ML_DF) # start 11:13

PA_ML_DF <- cbind(PA_ML_DF, PA_ML_DF1)
PA_ML_DF1 <- PA_ML_DF %>% drop_na(PCL_SE1) 

st_write(PA_ML_DF1,  here("data/processed/Pseudo_absences_ML.shp"), append = F, quiet = T)
rm(PA_ML_DF1)
```

```{r}
rm(WPDPV1, WPDPV2, BLM_AIM, BLM_CP, blm_cp_s, ecoregion_bound, test, test1, spp_occurrences_reg_random)
```

```{r}
PA_ML_DF <- st_read(here("data/processed/Pseudo_absences_ML.shp"), quiet = T)
PA_ML_DF_LIST <- split(PA_ML_DF, ~binomial)

Sys.time()
remaining_records <- lapply(PA_ML_DF_LIST, percentile)
Sys.time()

ab <- map(remaining_records, ~ .x %>% 
  mutate(
    pres_records = nrow(filter(., occurrence == 1)),
    true_abs = nrow(filter(., occurrence == 0 & !is.na(PlotKey))),
    pseudo_abs = nrow(filter(., occurrence == 0 & is.na(PlotKey))),
    required_records = pres_records - true_abs # records required for balanced P:A
    )
  ) %>%
  map(~ .x %>% 
        filter(occurrence == 0 & is.na(PlotKey)) %>% 
        slice_sample(n = slice_min(., required_records, n = 1, with_ties = F) %>% 
                 pull(required_records)
               )) %>% 
  bind_rows() %>% 
  dplyr::select(-pres_records, -true_abs, -required_records, -pseudo_abs) %>% 
  group_by(binomial) %>% 
  mutate(no_record = n())

remaining_records <- remaining_records %>% 
  bind_rows() %>% 
  filter(occurrence == 1 | occurrence == 0 & !is.na(PlotKey)) %>% 
  bind_rows(., ab)

st_write(remaining_records,  here("data/processed/Pseudo_absences_ML.shp"), append = F, quiet = T)
```

```{r}
rm(PA_ML_DF_LIST, ab, PA_ML_DF, remaining_records, spp_occurrences)
rm(list = lsf.str())
```


In an SDM built from a RF, simulation studies have shown that a roughly equivalent number of Presence to Absence records achieves the best results. Further results show that sampling the pseudo-absences from a range of environmental variables will produce better results than random stratification throughout geographic space. 

Herein we determine what realized niche values exist for our species of interest. We will extract the cell number of each of our presence records. We will then re-scale all of our predictor variables, and extract the values associated with our presence for each taxon. We will calculate a mean of these cells, with 100 bootstraps, and then remove cells from the raster within the 95% confidence intervals of the mean from the re-scaled row sums. We will then randomly sample a number of pseudo-absence records equivalent to the number of presences - 0.2 (the prop of records from BLM true absences). 

In this script we will limit the number of records to analyzed to the random points we have already drawn in the above process. 

## REFERENCES

Barbet-Massin, M., Jiguet, F., Albert, C.H., Thuiller, W. "Selecting pseudo-absences for species distribution models: how, where and how many?" Methods in Ecology and Evolution 2012, 3, 327–338 

