---
title: "phenology_windows"
author: "steppe"
date: '2022-04-22'
output: html_document
---

We extensively investigate the spatial filtering of ecological interactions. We make a short and cursory glance at the temporal filters upon ecological interactions. 

```{r load libraries, results='hide', message=F, warning=F}
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(raster) # raster data
library(here)
library(lubridate)
library(phenesse)
set.seed(12)
source(here::here('scripts/phen_functions.R'))
```


```{r}
spp_occurrences_full1 <- st_read(here("data/raw/BIEN_occurrence_records_full2022-04-22.shp"), quiet = T)
spp_occurrences_phen <- st_read(here("data/raw/BIEN_occ_rec_PHEN2022-05-01.shp"), quiet = T)
```


```{r Import and Buffer Field Station, warning = F}

ecoregions <- sf::read_sf(here("data/raw/us_eco_l4/us_eco_l4_no_st.shp")) %>% 
  st_transform(5070)

poly_coords <- st_read(
  paste0(here(), '/data/processed/BIEN_polygons', '2022-03-31', '.shp'), 
  quiet = T) 
poly100 <- poly_coords[7,] %>% 
  st_transform(st_crs(ecoregions))
poly100_bb <- poly100 %>%  
  st_bbox()

ecoregions <- ecoregions %>% 
  st_crop(poly100_bb)
rmbl <- st_centroid(poly100) %>% 
  st_transform(32613) %>% 
  st_buffer(15000) %>% 
  st_transform(st_crs(ecoregions))

ggplot() +
  geom_sf(data = ecoregions, aes(fill = US_L4NAME)) +
  geom_sf(data = rmbl, fill = NA, color = 'black', lwd = 1, lty = 5)

rm(poly_coords, poly100, poly100_bb)
```

```{r intersect field station to ecoregion, warning = F}
adj_L4 <- st_intersection(rmbl, ecoregions) %>% 
  distinct(US_L4NAME, .keep_all = T) %>% 
  pull(US_L4CODE)
ecoregions_subalp <- ecoregions %>% 
  filter(US_L4CODE %in% adj_L4) %>% 
  dplyr::select(US_L4CODE, US_L4NAME)

spp_occurrences_full1 <- st_transform(spp_occurrences_full1, st_crs(ecoregions_subalp))
spp_target_ecoregions <- st_intersection(spp_occurrences_full1, ecoregions_subalp) %>% 
  rename('Species' = scrbb__,
         'Date' = dt_cllc) %>% 
  drop_na(Date)

ggplot() +
  geom_sf(data = ecoregions_subalp, aes(fill = US_L4NAME)) +
  geom_sf(data = rmbl, fill = NA, color = 'black', lwd = 1, lty = 5) 

spp_occurrences_phen <- st_transform(spp_occurrences_phen, st_crs(ecoregions_subalp))
spp_target_ecoregions_phen <- st_intersection(spp_occurrences_phen, ecoregions_subalp) %>% 
  rename('Species' = scrbb__,
         'Date' = dt_cllc) %>% 
  drop_na(Date)

rm(poly100_t_dist, ecoregions_subalp, poly100_bb, adj_L4)
```


```{r calculate common snow free dates, warning = F}

snow_files <- paste0(paste0(here(), '/data/raw/snow_cover/'), 
                      list.files(paste0(here(), '/data/raw/snow_cover'))
)

snow_crs <- st_crs(brick(snow_files[1]))
gothic <- st_centroid(rmbl) %>% 
  st_transform(st_crs(snow_crs))
  
lower_snow_bound <- lwr_snw_bnd(snow_files, gothic)

```

```{r calculate first event where snow present for 3 or more days, warning = F, message = F}

upper_snow_bound <- uppr_snw_bnd(snow_files, gothic)

rm(x_year_snow,i, snow_files, rmbl, snow_crs, gothic, lwr_snw_bnd, uppr_snw_bnd)
```

We here calculate weibull based estimates of the date of flowering onset, cessation, and the dates when 10% of individuals are flowering, peak flowering (50% of individuals), and when 9-% of individuals have flowered. 

```{r calculate weibull phenology estimates 1}

spp_phen_recs <- bind_rows(spp_target_ecoregions_phen, spp_target_ecoregions)

cleaner_phenology_records <- spp_phen_recs %>% 
  dplyr::select(Species, Date) %>%
  group_by(Species) %>% 
  filter(!str_detect(Date,'\\d{4}-01-\\d{2}')) %>% # these are things without dates put in January
  mutate(DOY = lubridate::yday(as.Date(Date, format = "%Y-%m-%d")) + 1) %>%  # create doy column
  filter(DOY >= lower_snow_bound$third_quart[1], 
         DOY <= upper_snow_bound$first_quart[1]) %>%
  # only records after snow melt and before October are kept
  arrange(DOY) %>% 
  slice_head(prop = 0.9) %>% # remove some of the fruiting collections at the seasons end
  mutate(Observations  = n())

weibull_input <- cleaner_phenology_records %>% 
  filter(Observations >= 10) %>% # weibull requires 10 records to calculate.
  split(., .$Species)
```

```{r calculate weibull phenology estimates 2, echo = F, warning = F, eval = F}
weibull_estimates <- bind_rows(lapply(weibull_input, weibull_wrapper)) # about 3 hours for 300 spp
write.csv(weibull_estimates, paste0(here(),'/data/processed/weibull_estimates_phen.csv'),row.names = F)
```

```{r, message = F, warning=F}
rm(as, observations, today, weibull_input, weibull_wrapper, spp_target_ecoregions, spp_target_ecoregions_phen, spp_phen_recs)
```

We were able to estimate the above taxa using 10 or more records from purely sub-alpine settings. Here we glean more records from Montane forested settings. While this is not certainly not ideal as we are attempting to estimate the dates from which a tenth of individuals until 90% of individuals have flowered in the sub alpine valley of RMBL. We have no more records to use. 

```{r Additional mid-elevation forest records 1}

under_weibull_rec <- cleaner_phenology_records %>% 
  filter(Observations <= 9, Observations > 1) %>% 
  distinct(Species) %>% 
  pull(Species)

add_mnt_recs <- spp_occurrences_full1 %>% 
  filter(scrbb__ %in% under_weibull_rec)

ecoregions_mid_for <- ecoregions %>% 
  filter(US_L4NAME == 'Sedimentary Mid-Elevation Forests') %>% 
  dplyr::select(US_L4CODE, US_L4NAME)

add_mnt_recs <- st_transform(add_mnt_recs, st_crs(ecoregions_mid_for))
spp_target_montane <- st_intersection(add_mnt_recs, ecoregions_mid_for) %>% 
  rename('Species' = scrbb__,
       'Date' = dt_cllc) %>% 
  st_drop_geometry() %>% 
  drop_na(Date)

under_weibull_rec <- cleaner_phenology_records %>% 
  filter(Observations <= 9, Observations > 1) %>% 
  bind_rows(., spp_target_montane) %>% 
  dplyr::select(Species, Date) %>%
  group_by(Species) %>% 
  filter(!str_detect(Date,'\\d{4}-01-\\d{2}')) %>% # these are things without dates put in January
  mutate(DOY = lubridate::yday(as.Date(Date, format = "%Y-%m-%d")) + 1) %>%  # create doy column
  filter(DOY >= lower_snow_bound$third_quart[1], 
         DOY <= upper_snow_bound$first_quart[1]) %>% 
  arrange(DOY) %>% 
  mutate(Observations  = n()) %>% 
  st_drop_geometry()

weibull_input <- under_weibull_rec %>% 
  filter(Observations >= 10) %>% # weibull requires 10 records to calculate.
  split(., .$Species)
```

```{r Additional mid-elevation forest records 2, eval = F}
weibull_estimates <- bind_rows(lapply(weibull_input, weibull_wrapper)) 
write.csv(weibull_estimates, 
          paste0(here(),'/data/processed/weibull_estimates_mixed_montane.csv'),
          row.names = F)
```

```{r Additional mid-elevation forest records 3}
rm(add_mnt_recs, ecoregions_mid_for, weibull_input, ecoregions, ecoregions_mid_for,
   spp_target_montane, cleaner_phenology_records, spp_occurrences_full1)
```


```{r Species without enough records for weibull estimation, warning = F}

fixed_input <- under_weibull_rec %>% 
  filter(Observations <= 9, Observations > 1) %>% # take everything not suitable for weibull
  split(., .$Species)

fixed_estimates <- bind_rows(lapply(fixed_input, herb_wrapper)) %>% 
  mutate('Estimate' = 'Fixed-Montane')

rm(cleaner_phenology_records, herb_wrapper, fixed_input, upper_snow_bound, lower_snow_bound)
```

```{r Generate daily list of possible species in pollen loads}

weibull_estimates <- read.csv(paste0(here(),'/data/processed/weibull_estimates_phen.csv')) %>% 
  mutate('Estimate' = 'Weibull-Subalpine')
weibull_estimates1 <- read.csv(paste0(here(),'/data/processed/weibull_estimates_mixed_montane.csv')) %>% 
  mutate('Estimate' = 'Weibull-Montane')

flowering_est_long <- bind_rows(weibull_estimates, fixed_estimates) %>% 
  group_by(taxon) %>% 
  split(., .$taxon)

flowering_est_long <- bind_rows(lapply(flowering_est_long, daily_flowers)) %>% 
  arrange(DOY) %>% 
  group_by(DOY) 

flowering_est_long %>% 
  filter(DOY <= 244) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = DOY, y = count)) +
  geom_point()

rm(weibull_estimates, fixed_estimates, lwr_snw_bnd, upr_snw_bnd, daily_flowers, spp_occurrences_phen, under_weibull_rec, weibull_estimates1)
```


Belitz MW, Larsen EA, Ries L, Guralnick RP. The accuracy of phenology estimators for use with sparsely sampled presence-only observations. Methods Ecol Evol. 2020;11:1273–1285. https://doi.org/10.1111/2041-210X.13448

Tran, Hoang Viet; Nguyen, Phu; Ombadi, Mohammed; Hsu, Kuo-lin; Sorooshian, Soroosh; Xia, Qing (2018): A cloud-free MODIS snow cover dataset for the contiguous United States from 2000 to 2017. figshare. Dataset. https://doi.org/10.6084/m9.figshare.5902381.v4 

Iler, Amy. M, Humphrey, Parris T., Ogilvie,  Jane E., CaraDonna, Paul J. (2021) Conceptual and practical issues limit the utility of statistical estimators of phenological events. Ecosphere 2021, 12 (11). https://doi.org/10.1002/ecs2.3828
