---
title: "run_fit_SDMS_linear_models"
author: "steppe"
date: "3/10/2022"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    includes:
      in_header: float_free.tex
  word_document: default
---

\tableofcontents 
\listoffigures
\listoftables


This is simply the dedicated script where I run the modelling for SDM's for. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, results='hide', message=F, warning=F}
library(sdm) 
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(raster) # raster data
library(here)
source(here::here('scripts/sdm_functions.R'))
set.seed(12)
```

```{r Develop Raster Stack, echo = F}
names_vec <- c('PCL_MAN','PCL_SE1','PCL_SE2','PCL_SE3','PCM_BFP','PCM_CMD',
                  'PCM_DD5', 'PCM_MAP','PCM_PAS','PCM_TD','PCT_ECO','PCT_EFW',
                  'PLC_HRB', 'PLC_TRE', 'PLC_URB', 'PSL_BDR', 'PSL_CAR', 
                  'PSL_PHO', 'PSL_SND','PSL_TUS','PSL_TWL','PTP_ELV','PTP_RLV',
                  'PTP_SLP','PTP_WET','PTP_ASP','PTP_SOL', 'PCL_MRA','PSW_DIS',
                  'PSW_OCC','PCO_XSC', 'PCO_YSC')

WPDPV <- raster::stack("/hdd/Geospatial_data/Western_Plant_Predictors/SoRo_wppr_vers_two.tif")
names(WPDPV) <- names_vec
target <- c(2, 4,9, 10, 13, 14, 16, 19, 20, 23, 24, 25, 26, 27, 29, 30)

WPDPV2 <- WPDPV[[(target)]]
geo_proj = crs(WPDPV2)

rm(WPDPV, target, names_vec)
```

```{r Read in Species Occurrence Data, echo = F}
PA_REG_DF <- read_sf(here("data/processed/Pseudo_absences_regression.shp"))
PA_REG_DF %>% pull(binomial) %>% unique()
```

## D. Part 2: Run Linear Regression models, Generalized-Linear Model (GLM) and Generalized Additive Models (GAM), and ensemble results with adequate results.

Here we use the True Skill Statistic (TSS) statistic to develop our weighted ensemble from each of the predictions. We choose to use this has our primary diagnostic of model fit because it is able to work across a broad range of presence records. 

Allouche, O., Tsoar, A., Kadmon, K. "Assessing the accuracy of species distribution models: prevalence, kappa and the true skill statistic (TSS)" 2006. Journal of Applied Ecology 43: 1223-1232.  

```{r Run BLM/GAM based modelling, echo = F}

PA_REG_DF1 <- split(PA_REG_DF, f = PA_REG_DF$binomial)

Sys.time()
lapply(PA_REG_DF1[366:373], Linear_SDM) # skipped #3 & # 112
Sys.time()

```

