---
title: "species records rings"
author: "steppe"
date: "3/31/2022"
output: pdf_document
---

# Create a set of Polygons for searching the area around the study area to generate a species presence or absence list.

```{r load libraries, results='hide', message=F, warning=F}
library(tidyverse) # data tidying
library(sf) # spatial data compliant with tidyverse
library(here)
set.seed(12)
```

### A review of the basics geometry of an annulus and circle

Where the area of an annulus is calculated as:
$$
A = π(R^2 - r^2)
$$

In an example where the smallest radii of a circle is 10 km, and the annulus is 5 km.

$ A = π(15^2 - 10^2) = 138.23008 $

To determine the relationship between the area of the annulus and the circle within we may divide the area of the circle by that of that of the annulus.
$$
 \frac{A = π(r^2)}{A = π(R^2 - r^2)}
$$

$ 138.23008 / 314.2857  = 0.439823$

Our annulus in this example is then 0.43 times the size of the original circle.

### The numerical constant to createa circles doubled in size 

We want to double the area, and in order to do so, we’ll have to multiply the radius by some constant.
I will use *'d'* to represent the constant we want to *double the area* of a set of circles.

$2A = π(dr)^2$

$2A = πd^2r^2$

We may then replace A and the equivalent pi r squared
$2n = nd^2$

$2 = d^2$

$d = \sqrt 2 $

$d = 1.414214$

# Create many circle which double in radius

We will define the center of Gothic as such:
```{r}
Gothic <- tibble(
                 x = -106.98945,
                 y  = 38.95869
                 ) %>% 
  st_as_sf(coords = c(x = 'x', y = 'y'), 
           crs = 4326) %>% 
  st_transform(32613)
```

Our initial circle around the site will be 30 kilometers in diameter (18.6 miles), in my experience in an mountainous area this should return a *good many records of common* material.

Here we will also show the accuracy of circles drawn in a UTM system to deal with the geometric idea of a perfect circle
```{r determine accuracy of buffered distances, comment = F}

gothic1 <- Gothic %>% st_buffer(15000)

writeLines(paste0('A geometric circle is: ', 
       round((as.numeric(st_area(gothic1))) / (pi * 15000^2) * 100, 5),
' as large as a geographic one with planar coordinates.'))

rm(gothic1)
```

```{r Calculate values by hand to ensure accuracy, echo = F}
#round(15000*(sqrt(2)), 0) #= 21213 # first annulus
#round(21213*sqrt(2),0) #= 30000 # second annulus
#round(30000*sqrt(2),0) # = 42426 # third annulus
#round(42426*sqrt(2),0) # = 59999 # fourth annulus
#round(59999*sqrt(2),0) # = 84851 # fifth
#round(84851*sqrt(2),0) # = 119997 # sixth
```

```{r Calculate the radii required for equal area sampling, echo = F}
future_value = function(iterations, frst_radius = 1, calculation = 0.1) {
  frst_radius * calculation ^ (1:iterations)
}

radii = vector()
radii <- future_value(iterations = 6, frst_radius = 15000, calculation = sqrt(2))
radii2 <- tibble(radii) %>% 
  rownames_to_column('Polygon') %>% 
  mutate(Polygon = as.numeric(Polygon)+1) %>% 
  rename('Radius' = radii)
radius <- rbind(c('Polygon' = 1, radius = 15000), radii2)

rm(future_value, radii, radii2)
```


```{r Create polygons for sending to BIEN server}
poly_coords <- bind_cols(radius, Gothic) %>% 
  st_as_sf()
poly_coords <- st_buffer(poly_coords, poly_coords$Radius)

ggplot(poly_coords) +
  geom_sf(aes(fill = Polygon), alpha = 0.1)

poly_coords <- st_transform(poly_coords, 4326) 

rm(Gothic, radius)
```

```{r, eval = F}
st_write(poly_coords, 
         paste0(here(), '/data/processed/BIEN_polygons', Sys.Date(), '.shp'), 
         quiet = T)

rm(poly_coords)
```


